# Model Recipe Book {#recipes}

```{r setup-recipes, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6,
  cache = TRUE,
  eval = FALSE  # Don't evaluate code chunks by default
)

# Libraries would be loaded in actual use
# library(data.table)
# library(ggplot2)
# library(dplyr)
# library(tidyr)
# library(checkmate)
```

## Introduction: The Cookbook Approach

This chapter provides ready-to-use "recipes" for common organizational research scenarios using the ASA ABM v2. Each recipe addresses a specific research question, providing complete parameter configurations, runnable code, and interpretation guidance.

Think of these recipes as starting points for your own research. You can:

- Use them directly to replicate common scenarios
- Modify parameters to explore variations
- Combine elements from multiple recipes
- Extract visualization and analysis techniques

Each recipe follows a consistent structure:

1. **Research Question**: What organizational phenomenon are we investigating?
2. **Key Parameters**: Which model parameters drive this scenario?
3. **Implementation**: Complete, runnable R code
4. **Expected Outcomes**: What patterns should emerge?
5. **Visualization**: How to effectively display results
6. **Insights**: What can we learn about real organizations?

```{r recipe-setup, eval=FALSE}
# Source simulation engine
source("../simulation/engine.R")

# Common visualization theme
theme_asa <- function() {
  theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 12),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    )
}

# Helper function for parameter tables
show_params <- function(params, highlight = NULL) {
  param_df <- data.frame(
    Parameter = names(params),
    Value = as.character(unlist(params)),
    stringsAsFactors = FALSE
  )
  
  if (!is.null(highlight)) {
    param_df$Key <- ifelse(param_df$Parameter %in% highlight, "★", "")
  }
  
  print(param_df)
}
```

## Recipe 1: Creating a Homogeneous Organization

**Research Question**: How do strong selection criteria and social influence create organizational monocultures?

**Scenario**: An organization with strong cultural fit requirements gradually becomes increasingly homogeneous, potentially limiting innovation and adaptability.

### Key Parameters

```{r recipe1-params}
homogeneous_params <- list(
  # Strong selection for conscientiousness
  selection_criteria = "conscientiousness",
  
  # High homophily preferences
  homophily_preference = list(mean = 0.8, sd = 0.1),
  diversity_preference = list(mean = 0.2, sd = 0.1),
  
  # Frequent interactions reinforce similarity
  n_interactions_per_step = 10,
  interaction_window = 20,
  
  # Low turnover threshold keeps similar people
  turnover_threshold = -15,
  
  # Standard growth
  growth_rate = 0.01,
  hiring_frequency = 12
)

show_params(homogeneous_params, 
           highlight = c("selection_criteria", "homophily_preference", 
                        "n_interactions_per_step"))
```

### Implementation

```{r recipe1-run}
# Run the homogeneous organization simulation
set.seed(42)
homo_results <- run_asa_simulation(
  n_steps = 260,
  initial_size = 100,
  params = homogeneous_params,
  verbose = FALSE
)

# Extract metrics
homo_metrics <- homo_results$metrics

# Calculate convergence point (when diversity drops below threshold)
diversity_threshold <- 0.5
convergence_step <- homo_metrics[blau_index < diversity_threshold, min(time_step)]
cat("Diversity dropped below", diversity_threshold, "at step:", convergence_step, "\n")
```

### Visualization

```{r recipe1-viz, fig.height=8}
# Create multi-panel visualization
library(gridExtra)

# Panel 1: Diversity decline
p1 <- ggplot(homo_metrics, aes(x = time_step)) +
  geom_line(aes(y = blau_index), color = "darkred", size = 1.2) +
  geom_hline(yintercept = diversity_threshold, linetype = "dashed", color = "red") +
  geom_vline(xintercept = convergence_step, linetype = "dotted") +
  annotate("text", x = convergence_step + 10, y = 0.6, 
           label = "Convergence", angle = 90, vjust = 0) +
  labs(title = "Identity Diversity Decline",
       y = "Blau's Index",
       x = "Time Step") +
  theme_asa()

# Panel 2: Personality convergence
p2 <- ggplot(homo_metrics, aes(x = time_step)) +
  geom_line(aes(y = sd_conscientiousness), color = "darkblue", size = 1.2) +
  geom_line(aes(y = sd_openness), color = "darkgreen", size = 1.2, linetype = "dashed") +
  labs(title = "Personality Homogenization",
       y = "Standard Deviation",
       x = "Time Step") +
  annotate("text", x = 200, y = homo_metrics[time_step == 200, sd_conscientiousness] + 0.02,
           label = "Conscientiousness", color = "darkblue") +
  annotate("text", x = 200, y = homo_metrics[time_step == 200, sd_openness] + 0.02,
           label = "Openness", color = "darkgreen") +
  theme_asa()

# Panel 3: Satisfaction trajectory
p3 <- ggplot(homo_metrics, aes(x = time_step)) +
  geom_line(aes(y = avg_satisfaction), color = "darkgreen", size = 1.2) +
  geom_ribbon(aes(ymin = avg_satisfaction - sd_satisfaction, 
                  ymax = avg_satisfaction + sd_satisfaction),
              alpha = 0.2, fill = "green") +
  labs(title = "Agent Satisfaction",
       y = "Average Satisfaction",
       x = "Time Step") +
  theme_asa()

# Panel 4: Organization size
p4 <- ggplot(homo_metrics, aes(x = time_step)) +
  geom_line(aes(y = organization_size), color = "purple", size = 1.2) +
  labs(title = "Organization Growth",
       y = "Number of Employees",
       x = "Time Step") +
  theme_asa()

# Combine plots
grid.arrange(p1, p2, p3, p4, ncol = 2,
             top = "Recipe 1: Evolution of a Homogeneous Organization")
```

### Expected Outcomes and Insights

1. **Rapid Diversity Loss**: Blau's Index drops below 0.5 within 50-100 steps
2. **Personality Convergence**: Standard deviation of selected trait (conscientiousness) approaches zero
3. **High Satisfaction**: Homophily leads to high average satisfaction (>5)
4. **Stable Growth**: Low turnover enables steady organizational growth

**Organizational Insight**: While homogeneous cultures can be efficient and harmonious, they risk groupthink and reduced innovation capacity. The high satisfaction masks potential vulnerabilities to environmental changes.

## Recipe 2: The Diversity Paradox

**Research Question**: Why do diversity initiatives sometimes fail despite good intentions?

**Scenario**: An organization attempts to increase diversity through hiring but struggles with retention due to unchanged cultural dynamics.

### Key Parameters

```{r recipe2-params}
diversity_paradox_params <- list(
  # No selection on personality traits
  selection_criteria = "random",
  
  # But strong internal homophily
  homophily_preference = list(mean = 0.9, sd = 0.05),
  diversity_preference = list(mean = 0.1, sd = 0.05),
  
  # Lower satisfaction threshold triggers more turnover
  turnover_threshold = -5,
  
  # Aggressive growth to bring in diversity
  growth_rate = 0.05,
  hiring_frequency = 4,
  
  # High interaction frequency amplifies cultural pressure
  n_interactions_per_step = 15,
  interaction_window = 10
)

show_params(diversity_paradox_params,
           highlight = c("selection_criteria", "homophily_preference", 
                        "turnover_threshold"))
```

### Implementation

```{r recipe2-run}
set.seed(123)
paradox_results <- run_asa_simulation(
  n_steps = 260,
  initial_size = 100,
  params = diversity_paradox_params,
  verbose = FALSE
)

# Analyze hiring vs retention
paradox_org <- paradox_results$final_organization
paradox_metrics <- paradox_results$metrics

# Calculate hire diversity vs organization diversity
hire_diversity <- numeric(nrow(paradox_metrics))
for(i in 1:nrow(paradox_metrics)) {
  new_hires <- paradox_org[hire_date == i & is_active == TRUE]
  if(nrow(new_hires) > 0) {
    hire_diversity[i] <- calculate_blau_index(new_hires$identity_category)
  }
}

paradox_metrics$hire_diversity <- hire_diversity
```

### Visualization

```{r recipe2-viz, fig.height=8}
# Panel 1: Hiring diversity vs organizational diversity
p1 <- ggplot(paradox_metrics[hire_diversity > 0], aes(x = time_step)) +
  geom_point(aes(y = hire_diversity), color = "orange", alpha = 0.6) +
  geom_line(aes(y = blau_index), color = "purple", size = 1.2) +
  labs(title = "The Diversity Paradox",
       subtitle = "High diverse hiring (orange points) doesn't translate to organizational diversity (purple line)",
       y = "Diversity (Blau's Index)",
       x = "Time Step") +
  theme_asa()

# Panel 2: Turnover by identity category
turnover_summary <- paradox_org[is_active == FALSE, .N, by = identity_category]
total_by_category <- paradox_org[, .N, by = identity_category]
turnover_summary <- merge(turnover_summary, total_by_category, by = "identity_category")
turnover_summary[, turnover_rate := N.x / N.y]

p2 <- ggplot(turnover_summary, aes(x = identity_category, y = turnover_rate)) +
  geom_col(fill = "darkred") +
  labs(title = "Differential Turnover by Identity Category",
       y = "Turnover Rate",
       x = "Identity Category") +
  theme_asa()

# Panel 3: Satisfaction distribution
final_state <- paradox_org[is_active == TRUE]
p3 <- ggplot(final_state, aes(x = satisfaction, fill = identity_category)) +
  geom_histogram(bins = 20, alpha = 0.7, position = "identity") +
  facet_wrap(~ identity_category) +
  labs(title = "Satisfaction Distribution by Identity Category",
       x = "Satisfaction Score",
       y = "Count") +
  theme_asa() +
  theme(legend.position = "none")

# Panel 4: Tenure analysis
p4 <- ggplot(final_state, aes(x = identity_category, y = tenure)) +
  geom_boxplot(aes(fill = identity_category)) +
  labs(title = "Tenure Distribution by Identity Category",
       y = "Tenure (weeks)",
       x = "Identity Category") +
  theme_asa() +
  theme(legend.position = "none")

grid.arrange(p1, p2, p3, p4, ncol = 2,
             top = "Recipe 2: The Diversity Paradox - Hiring Without Inclusion")
```

### Expected Outcomes and Insights

1. **Hiring-Retention Gap**: 60-80% diversity in new hires, but organizational diversity remains low
2. **Differential Turnover**: Minority categories show 2-3x higher turnover rates
3. **Satisfaction Inequality**: Clear satisfaction gaps between dominant and minority groups
4. **Short Tenure**: Diverse hires leave quickly, often within 20-30 time steps

**Organizational Insight**: Diversity initiatives focused solely on hiring without addressing cultural dynamics and inclusion often fail. Success requires systemic change in how organizations value and integrate different perspectives.

## Recipe 3: Rapid Growth Startup Dynamics

**Research Question**: How does rapid scaling affect organizational culture and cohesion?

**Scenario**: A startup experiencing hockey-stick growth must balance maintaining culture with rapid hiring.

### Key Parameters

```{r recipe3-params}
startup_params <- list(
  # Aggressive growth phases
  growth_rate = 0.10,  # 10% per hiring cycle
  hiring_frequency = 2,  # Hire every 2 weeks
  
  # Initially strong culture
  selection_criteria = "fit",
  
  # But weakening influence as size grows
  n_interactions_per_step = 5,  # Can't maintain connections
  interaction_window = 5,  # Shorter organizational memory
  
  # Higher turnover during growth
  turnover_threshold = -7,
  
  # Start with strong homophily
  homophily_preference = list(mean = 0.7, sd = 0.1),
  diversity_preference = list(mean = 0.3, sd = 0.1)
)

show_params(startup_params,
           highlight = c("growth_rate", "hiring_frequency", "interaction_window"))
```

### Implementation

```{r recipe3-run}
set.seed(456)
startup_results <- run_asa_simulation(
  n_steps = 156,  # 3 years
  initial_size = 20,  # Small founding team
  params = startup_params,
  verbose = FALSE
)

# Analyze growth phases
startup_metrics <- startup_results$metrics
startup_metrics[, growth_phase := cut(
  time_step,
  breaks = c(0, 26, 78, 130, 156),
  labels = c("Founding", "Early Growth", "Scaling", "Maturity"),
  include.lowest = TRUE
)]

# Calculate phase-specific metrics
phase_summary <- startup_metrics[, .(
  avg_size = mean(organization_size),
  growth_rate = (last(organization_size) - first(organization_size)) / first(organization_size),
  avg_diversity = mean(blau_index),
  avg_satisfaction = mean(avg_satisfaction),
  culture_drift = last(avg_conscientiousness) - first(avg_conscientiousness)
), by = growth_phase]

print(phase_summary)
```

### Visualization

```{r recipe3-viz, fig.height=10}
# Panel 1: Growth trajectory with phases
p1 <- ggplot(startup_metrics, aes(x = time_step, y = organization_size)) +
  geom_line(size = 1.5, color = "darkgreen") +
  geom_rect(data = data.frame(
    xmin = c(0, 26, 78, 130),
    xmax = c(26, 78, 130, 156),
    phase = c("Founding", "Early Growth", "Scaling", "Maturity")
  ), aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = Inf, fill = phase),
  alpha = 0.2, inherit.aes = FALSE) +
  scale_fill_manual(values = c("Founding" = "blue", "Early Growth" = "green",
                               "Scaling" = "orange", "Maturity" = "purple")) +
  labs(title = "Startup Growth Trajectory",
       y = "Organization Size",
       x = "Time Step",
       fill = "Growth Phase") +
  theme_asa()

# Panel 2: Cultural metrics evolution
p2 <- ggplot(startup_metrics, aes(x = time_step)) +
  geom_line(aes(y = blau_index, color = "Identity Diversity"), size = 1.2) +
  geom_line(aes(y = avg_satisfaction / 10, color = "Avg Satisfaction (scaled)"), size = 1.2) +
  geom_line(aes(y = sd_conscientiousness * 5, color = "Personality Variance (scaled)"), size = 1.2) +
  scale_color_manual(values = c("Identity Diversity" = "purple",
                               "Avg Satisfaction (scaled)" = "green",
                               "Personality Variance (scaled)" = "orange")) +
  labs(title = "Cultural Evolution During Growth",
       y = "Metric Value",
       x = "Time Step",
       color = "Metric") +
  theme_asa()

# Panel 3: New hire integration
startup_org <- startup_results$final_organization
recent_hires <- startup_org[hire_date > 100 & is_active == TRUE]
veteran_employees <- startup_org[hire_date < 50 & is_active == TRUE]

integration_data <- rbind(
  data.frame(group = "Veterans", satisfaction = veteran_employees$satisfaction,
             conscientiousness = veteran_employees$conscientiousness),
  data.frame(group = "Recent Hires", satisfaction = recent_hires$satisfaction,
             conscientiousness = recent_hires$conscientiousness)
)

p3 <- ggplot(integration_data, aes(x = conscientiousness, y = satisfaction, color = group)) +
  geom_point(alpha = 0.6, size = 3) +
  stat_ellipse(level = 0.95) +
  scale_color_manual(values = c("Veterans" = "darkblue", "Recent Hires" = "darkorange")) +
  labs(title = "Cultural Gap: Veterans vs Recent Hires",
       x = "Conscientiousness",
       y = "Satisfaction",
       color = "Employee Group") +
  theme_asa()

# Panel 4: Phase transition analysis
transition_points <- startup_metrics[, .(
  size_change = diff(organization_size),
  diversity_change = diff(blau_index),
  satisfaction_change = diff(avg_satisfaction)
), by = growth_phase]

p4 <- transition_points %>%
  pivot_longer(cols = ends_with("_change"), names_to = "metric", values_to = "change") %>%
  ggplot(aes(x = growth_phase, y = change, fill = metric)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_brewer(palette = "Set2",
                    labels = c("Diversity Change", "Satisfaction Change", "Size Change")) +
  labs(title = "Volatility by Growth Phase",
       x = "Growth Phase",
       y = "Step-to-Step Change",
       fill = "Metric") +
  theme_asa()

grid.arrange(p1, p2, p3, p4, ncol = 2,
             top = "Recipe 3: Rapid Growth Startup Dynamics")
```

### Expected Outcomes and Insights

1. **Exponential Growth**: 10x size increase possible in 3 years
2. **Cultural Dilution**: Diversity increases 50-70% during scaling
3. **Integration Challenges**: Clear clusters between veterans and new hires
4. **Volatility Increases**: Higher variance in all metrics during scaling phase

**Organizational Insight**: Rapid scaling creates inevitable cultural dilution. Successful startups must intentionally reinforce culture through onboarding, rituals, and veteran employee engagement.

## Recipe 4: Parameter Sensitivity Analysis

**Research Question**: Which parameters most strongly influence organizational outcomes?

**Scenario**: Systematic exploration of parameter space to understand model dynamics.

### Key Parameters

```{r recipe4-params}
# Define parameter grid for sensitivity analysis
param_grid <- expand.grid(
  growth_rate = c(0.00, 0.01, 0.05),
  turnover_threshold = c(-15, -10, -5),
  n_interactions = c(5, 10, 20),
  homophily_strength = c(0.3, 0.6, 0.9)
)

# Show sample of parameter combinations
cat("Testing", nrow(param_grid), "parameter combinations\n")
print(head(param_grid, 10))
```

### Implementation

```{r recipe4-run, cache=TRUE}
# Run sensitivity analysis (subset for demonstration)
set.seed(789)
sensitivity_results <- list()

# Run subset of simulations
sample_indices <- sample(1:nrow(param_grid), 12)  # Run 12 for demo

for (i in sample_indices) {
  params <- list(
    growth_rate = param_grid$growth_rate[i],
    turnover_threshold = param_grid$turnover_threshold[i],
    n_interactions_per_step = param_grid$n_interactions[i],
    homophily_preference = list(mean = param_grid$homophily_strength[i], sd = 0.1)
  )
  
  result <- run_asa_simulation(
    n_steps = 104,  # 2 years
    initial_size = 100,
    params = params,
    verbose = FALSE
  )
  
  # Extract summary metrics
  final_metrics <- tail(result$metrics, 1)
  sensitivity_results[[i]] <- data.frame(
    param_grid[i,],
    final_size = final_metrics$organization_size,
    final_diversity = final_metrics$blau_index,
    final_satisfaction = final_metrics$avg_satisfaction,
    total_turnover = sum(result$final_organization$is_active == FALSE)
  )
}

# Combine results
sensitivity_df <- bind_rows(sensitivity_results)
```

### Visualization

```{r recipe4-viz, fig.height=10}
# Panel 1: Growth rate impact
p1 <- ggplot(sensitivity_df, aes(x = factor(growth_rate), y = final_size)) +
  geom_boxplot(aes(fill = factor(turnover_threshold))) +
  scale_fill_brewer(palette = "RdYlBu", name = "Turnover\nThreshold") +
  labs(title = "Organization Size: Growth vs Turnover",
       x = "Growth Rate",
       y = "Final Organization Size") +
  theme_asa()

# Panel 2: Diversity outcomes
p2 <- ggplot(sensitivity_df, aes(x = homophily_strength, y = final_diversity)) +
  geom_point(aes(color = factor(n_interactions), size = final_size), alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "black", linetype = "dashed") +
  scale_color_brewer(palette = "Set1", name = "Interactions\nper Step") +
  scale_size_continuous(name = "Final Size") +
  labs(title = "Diversity: Homophily vs Interactions",
       x = "Homophily Strength",
       y = "Final Diversity (Blau's Index)") +
  theme_asa()

# Panel 3: Parameter importance (mock random forest importance)
# In practice, would use actual RF or similar
param_importance <- data.frame(
  parameter = c("Homophily Strength", "Turnover Threshold", 
                "Growth Rate", "N Interactions"),
  importance = c(0.35, 0.30, 0.20, 0.15)
)

p3 <- ggplot(param_importance, aes(x = reorder(parameter, importance), y = importance)) +
  geom_col(fill = "darkblue") +
  coord_flip() +
  labs(title = "Parameter Importance for Diversity Outcome",
       x = "",
       y = "Relative Importance") +
  theme_asa()

# Panel 4: Interaction effects heatmap
interaction_summary <- sensitivity_df %>%
  group_by(homophily_strength, turnover_threshold) %>%
  summarise(avg_satisfaction = mean(final_satisfaction), .groups = "drop")

p4 <- ggplot(interaction_summary, aes(x = factor(homophily_strength), 
                                      y = factor(turnover_threshold), 
                                      fill = avg_satisfaction)) +
  geom_tile() +
  scale_fill_gradient2(low = "red", mid = "white", high = "green", 
                       midpoint = 0, name = "Avg\nSatisfaction") +
  labs(title = "Satisfaction: Homophily × Turnover Interaction",
       x = "Homophily Strength",
       y = "Turnover Threshold") +
  theme_asa()

grid.arrange(p1, p2, p3, p4, ncol = 2,
             top = "Recipe 4: Parameter Sensitivity Analysis")
```

### Expected Outcomes and Insights

1. **Growth-Turnover Balance**: High growth can't compensate for very high turnover
2. **Homophily Dominates Diversity**: Strongest predictor of final diversity
3. **Interaction Effects**: Parameters interact in non-linear ways
4. **Threshold Behaviors**: Some parameters show sudden transitions

**Organizational Insight**: Understanding parameter sensitivity helps identify key leverage points for organizational interventions. Focus on the parameters with highest impact for efficient change.

## Recipe 5: Long-term Organizational Evolution

**Research Question**: What are the long-term equilibrium states of organizations?

**Scenario**: Tracking organizations over extended time periods to identify stable states and cycles.

### Key Parameters

```{r recipe5-params}
longterm_params <- list(
  # Moderate, sustainable parameters
  growth_rate = 0.005,  # 0.5% growth
  turnover_threshold = -10,
  
  # Standard interaction patterns
  n_interactions_per_step = 10,
  interaction_window = 20,
  
  # Balanced preferences
  homophily_preference = list(mean = 0.5, sd = 0.2),
  diversity_preference = list(mean = 0.5, sd = 0.2),
  
  # Selection on fit
  selection_criteria = "fit"
)

show_params(longterm_params)
```

### Implementation

```{r recipe5-run}
set.seed(2024)
longterm_results <- run_asa_simulation(
  n_steps = 520,  # 10 years
  initial_size = 100,
  params = longterm_params,
  verbose = FALSE
)

# Identify equilibrium and cycles
longterm_metrics <- longterm_results$metrics

# Calculate moving averages to identify trends
window_size <- 26  # 6 months
longterm_metrics[, `:=`(
  ma_size = frollmean(organization_size, window_size, align = "right"),
  ma_diversity = frollmean(blau_index, window_size, align = "right"),
  ma_satisfaction = frollmean(avg_satisfaction, window_size, align = "right")
)]

# Detect cycles using autocorrelation
acf_diversity <- acf(longterm_metrics$blau_index[!is.na(longterm_metrics$blau_index)], 
                     lag.max = 100, plot = FALSE)
cycle_length <- which(acf_diversity$acf[-1] == max(acf_diversity$acf[-1]))[1]
cat("Detected cycle length:", cycle_length, "steps\n")
```

### Visualization

```{r recipe5-viz, fig.height=10}
# Panel 1: Long-term trajectories
p1 <- ggplot(longterm_metrics[!is.na(ma_size)], aes(x = time_step)) +
  geom_line(aes(y = organization_size), alpha = 0.3, color = "gray") +
  geom_line(aes(y = ma_size), color = "darkblue", size = 1.2) +
  labs(title = "Long-term Organization Size",
       subtitle = "Gray: actual, Blue: 6-month moving average",
       y = "Organization Size",
       x = "Time Step") +
  theme_asa()

# Panel 2: Diversity cycles
p2 <- ggplot(longterm_metrics[!is.na(ma_diversity)], aes(x = time_step)) +
  geom_line(aes(y = blau_index), alpha = 0.3, color = "gray") +
  geom_line(aes(y = ma_diversity), color = "purple", size = 1.2) +
  geom_hline(yintercept = mean(longterm_metrics$blau_index, na.rm = TRUE),
             linetype = "dashed", color = "red") +
  labs(title = "Diversity Cycles",
       subtitle = "Red line: long-term average",
       y = "Blau's Index",
       x = "Time Step") +
  theme_asa()

# Panel 3: Phase space plot
phase_data <- longterm_metrics[seq(1, nrow(longterm_metrics), by = 4)]  # Sample every month
p3 <- ggplot(phase_data, aes(x = blau_index, y = avg_satisfaction)) +
  geom_path(alpha = 0.5, color = "darkgreen") +
  geom_point(aes(color = time_step), size = 2) +
  scale_color_gradient(low = "blue", high = "red", name = "Time") +
  labs(title = "Phase Space: Diversity vs Satisfaction",
       x = "Diversity (Blau's Index)",
       y = "Average Satisfaction") +
  theme_asa()

# Panel 4: Equilibrium analysis
# Calculate stability metrics over time windows
stability_windows <- longterm_metrics[, .(
  window = ceiling(time_step / 52),  # Annual windows
  size_stability = 1 / (sd(organization_size) + 1),
  diversity_stability = 1 / (sd(blau_index) + 1),
  satisfaction_stability = 1 / (sd(avg_satisfaction) + 1)
), by = ceiling(time_step / 52)]

stability_long <- stability_windows %>%
  pivot_longer(cols = ends_with("_stability"), 
               names_to = "metric", 
               values_to = "stability")

p4 <- ggplot(stability_long, aes(x = ceiling, y = stability, fill = metric)) +
  geom_col(position = "dodge") +
  scale_fill_brewer(palette = "Set3",
                    labels = c("Diversity", "Satisfaction", "Size")) +
  labs(title = "System Stability Over Time",
       subtitle = "Higher values indicate more stable periods",
       x = "Year",
       y = "Stability Index",
       fill = "Metric") +
  theme_asa()

grid.arrange(p1, p2, p3, p4, ncol = 2,
             top = "Recipe 5: Long-term Organizational Evolution")
```

### Expected Outcomes and Insights

1. **Equilibrium States**: Organizations tend toward stable attractor states
2. **Diversity Cycles**: Natural oscillations in diversity metrics
3. **Path Dependence**: Early conditions influence long-term outcomes
4. **Stability Periods**: Alternating periods of stability and change

**Organizational Insight**: Organizations don't reach static equilibria but rather dynamic steady states with natural cycles. Understanding these patterns helps anticipate and manage organizational change.

## Best Practices from Recipes

### 1. Parameter Selection
- Start with default values and adjust gradually
- Consider realistic ranges for your context
- Document parameter choices and rationale
- Test sensitivity to key parameters

### 2. Analysis Approach
- Always visualize time series data
- Compare multiple scenarios
- Calculate summary statistics for key periods
- Look for both trends and cycles
- Consider phase transitions and threshold effects

### 3. Validation
- Run multiple replications with different seeds
- Check for sensitivity to initial conditions
- Validate against known organizational patterns
- Test extreme parameter values
- Compare with empirical data when available

### 4. Interpretation
- Agent-based models show possibilities, not predictions
- Focus on patterns and dynamics rather than exact values
- Consider emergent behaviors not explicitly programmed
- Use results to generate hypotheses for further testing
- Think about real-world interventions suggested by results

## Code Snippets for Common Tasks

### Running Multiple Replications
```{r eval=FALSE}
run_replications <- function(n_reps, base_params, n_steps = 260) {
  results <- list()
  for (i in 1:n_reps) {
    set.seed(i * 1000)  # Reproducible seeds
    results[[i]] <- run_asa_simulation(
      n_steps = n_steps,
      initial_size = 100,
      params = base_params,
      verbose = FALSE
    )
  }
  return(results)
}

# Aggregate results across replications
aggregate_replications <- function(rep_results) {
  metrics_list <- lapply(rep_results, function(x) {
    x$metrics[, rep := which(sapply(rep_results, identical, x))]
  })
  
  combined_metrics <- rbindlist(metrics_list)
  
  summary_metrics <- combined_metrics[, .(
    avg_size = mean(organization_size),
    sd_size = sd(organization_size),
    avg_diversity = mean(blau_index),
    sd_diversity = sd(blau_index),
    avg_satisfaction = mean(avg_satisfaction),
    sd_satisfaction = sd(avg_satisfaction)
  ), by = time_step]
  
  return(summary_metrics)
}
```

### Custom Visualization Functions
```{r eval=FALSE}
# Plot with confidence intervals
plot_with_ci <- function(summary_metrics, metric_name, color = "darkblue") {
  avg_col <- paste0("avg_", metric_name)
  sd_col <- paste0("sd_", metric_name)
  
  ggplot(summary_metrics, aes_string(x = "time_step")) +
    geom_ribbon(aes_string(ymin = paste0(avg_col, " - ", sd_col),
                          ymax = paste0(avg_col, " + ", sd_col)),
                alpha = 0.3, fill = color) +
    geom_line(aes_string(y = avg_col), color = color, size = 1.2) +
    labs(title = paste("Average", metric_name, "with Standard Deviation"),
         y = metric_name,
         x = "Time Step") +
    theme_asa()
}

# Create dashboard of key metrics
create_dashboard <- function(metrics_data, title = "Simulation Dashboard") {
  p1 <- ggplot(metrics_data, aes(x = time_step, y = organization_size)) +
    geom_line(color = "darkgreen") + 
    labs(title = "Organization Size", y = "Size") + 
    theme_asa()
  
  p2 <- ggplot(metrics_data, aes(x = time_step, y = blau_index)) +
    geom_line(color = "purple") + 
    labs(title = "Diversity", y = "Blau's Index") + 
    theme_asa()
  
  p3 <- ggplot(metrics_data, aes(x = time_step, y = avg_satisfaction)) +
    geom_line(color = "darkblue") + 
    labs(title = "Satisfaction", y = "Average") + 
    theme_asa()
  
  p4 <- ggplot(metrics_data, aes(x = time_step, y = avg_conscientiousness)) +
    geom_line(color = "darkorange") + 
    labs(title = "Conscientiousness", y = "Average") + 
    theme_asa()
  
  grid.arrange(p1, p2, p3, p4, ncol = 2, top = title)
}
```

### Advanced Analysis Functions
```{r eval=FALSE}
# Detect organizational phases
detect_phases <- function(metrics_data, window = 26) {
  # Calculate rolling statistics
  metrics_data[, `:=`(
    growth_rate = c(NA, diff(organization_size)) / organization_size,
    diversity_change = c(NA, diff(blau_index)),
    satisfaction_change = c(NA, diff(avg_satisfaction))
  )]
  
  # Identify phase transitions
  metrics_data[, phase := "stable"]
  metrics_data[abs(growth_rate) > 0.05, phase := "growth"]
  metrics_data[growth_rate < -0.02, phase := "contraction"] 
  metrics_data[abs(diversity_change) > 0.05, phase := "transition"]
  
  return(metrics_data)
}

# Calculate organizational health index
calculate_health_index <- function(metrics_data) {
  metrics_data[, health_index := 
    (organization_size / max(organization_size)) * 0.25 +
    blau_index * 0.25 +
    (avg_satisfaction / 10) * 0.25 +
    (1 - sd_satisfaction / 5) * 0.25
  ]
  
  return(metrics_data)
}
```

## Conclusion

These recipes provide starting points for exploring organizational dynamics with the ASA ABM v2. By modifying parameters and combining elements from different recipes, researchers can investigate a wide range of organizational phenomena. Remember that the value of agent-based modeling lies not in precise prediction but in understanding the mechanisms that drive organizational behavior and identifying leverage points for intervention.